---
title: "Developmental milestone curves from parent report data"
author: "Mike & Emily"
date: "`r Sys.Date()`"
output: 
html_document:
toc: true
number_sections: true
runtime: shiny
---

```{r, echo=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=5, 
                      echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE)
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(langcog))
library(stringr)
library(tidyr)
library(ggrepel)
library(binom)
library(shiny)
library(readr)
theme_set(theme_bw())
```

```{r, echo=FALSE}
solarized_colors <- c("#268bd2", "#cb4b16", "#859900", "#993399", "#d33682",
                      "#b58900", "#2aa198", "#6c71c4", "#dc322f")

stable_order_palete <- function(num_values) {
  c(rep(solarized_colors, num_values %/% length(solarized_colors)),
    solarized_colors[1:(num_values %% length(solarized_colors))])
}

# minimum N for inclusions of a datapoint
MIN_N <- 50
```

# Introduction

We are analyzing data from Kinedu, an app for parents of young children to do activities with their children. 

```{r}
load("data/kinedu_all_processed_3-11-16.RData")
info <- read_csv("info.csv") %>%
  filter(!duplicated(name))

ms <- right_join(ms, info) %>% # remove duplicated indicators
  select(indicator_id, age_months, answer, n, general_category, 
         specific_category, name, cat_name, starts_with("WHO"), starts_with("CDC"), redundant) 
```

The particular data we look at here are a set of `r prettyNum(sum(ms$n), big.mark=",", scientific=FALSE)` binary answers given the first time parents saw a set of `r length(unique(ms$name))` "indicators" - questions that the app asked to help with developmental categorization. 

Some (random) example indicators:

+ "`r unique(ms$name)[20]`," in the category "`r ms$specific_category[ms$name == unique(ms$name)[20]][1]`"
+ "`r unique(ms$name)[40]`," in the category "`r ms$specific_category[ms$name == unique(ms$name)[40]][1]`"
+ "`r unique(ms$name)[60]`," in the category "`r ms$specific_category[ms$name == unique(ms$name)[60]][1]`"

This report uses an (in-progress) manual categorization of markers to examine their developmental curves, both individually and in aggregate.

# Major milestones

Here is a set of major milestones of interest. 

```{r}
plot_data <- filter(ms, cat_name %in% c("cognitive","crawl","smile","walk"))
selected_descs <- c("Begins to walk alone", 
                    "Begins to crawl",
                    "Smiles and begins to babble",
                    "Says one to four words with meaning ",
                    "Switches objects from one hand to the other")

plot_data <- filter(ms, 
                    name %in% selected_descs) %>%
  filter(n > MIN_N) %>%
  mutate(name = factor(name, levels = selected_descs), 
         age_years = age_months / 12, 
         ci.low = binom::binom.confint(x = answer * n, n = n, 
                                       methods = "bayes")$lower,
         ci.high = binom::binom.confint(x = answer * n, n = n, 
                                        methods = "bayes")$upper)

label_data <- plot_data %>% 
  group_by(name) %>% 
  summarise(answer = median(answer), 
            age_years = median(age_years))

ggplot(plot_data,
       aes(x = age_years, y = answer, col = name)) + 
  geom_point(aes(size = n)) + 
  geom_linerange(aes(ymin = ci.low, ymax = ci.high)) + 
  # geom_smooth(se=FALSE, aes(weight = n), method = "loess") + 
  geom_line() + 
  scale_colour_manual(guide=FALSE, 
                      values = stable_order_palete(length(unique(selected_descs)))) +
  scale_fill_manual(guide=FALSE, 
                    values = stable_order_palete(length(unique(selected_descs)))) + 
  scale_size_continuous(guide= FALSE) + 
  ylim(c(0,1)) + 
  xlim(c(0,2)) +
  xlab("Age (years)") + 
  ylab("Proportion parents responding yes") + 
  geom_label_repel(data = label_data, aes(fill = name, 
                                          label = name), 
                   force = 8, size = 3,
                   color = 'white',
                   box.padding = unit(0.25, "lines"),
                   point.padding = unit(0.5, "lines")) 
```

Size of dots shows the number of datapoints, and vertical lines show 95% confidence intervals. Only ages with more than 50 observations are shown.

# Interactive exploration

Select a general category, a specific category, and then milestones from that category. 

```{r, cache=FALSE}
shinyApp(
  ui = fluidPage(
    sidebarLayout(
      sidebarPanel(
        selectInput("gen_cat", "General Category", 
                    selected = "touch", 
                    choices = as.character(unique(ms$general_category))),
        uiOutput("spec_selector"),
        uiOutput("desc_selector")
      ),
      mainPanel(
        plotOutput("catplot"))
    )
  ),
  
  server = function(input, output) {
    
    specs <- reactive({
      as.character(unique(ms$specific_category[ms$general_category == input$gen_cat]))
    })
    
    descs <- reactive({
      as.character(unique(ms$name[ms$general_category == input$gen_cat &
                                    ms$specific_category == input$spec_cat]))
    })
    
    output$spec_selector <- renderUI({
      selectInput("spec_cat", "Specific Category", 
                  choices = specs(), 
                  selected = specs()[1])
    })
    
    output$desc_selector <- renderUI({
      selectizeInput("selected_descs", "Labels", 
                     choices = descs(), 
                     selected = descs()[1],
                     options = list(maxItems = 9), 
                     multiple = TRUE)
    })
    
    output$catplot <- renderPlot({
      
      plot_data <- filter(ms, 
                          specific_category == input$spec_cat, 
                          general_category == input$gen_cat,
                          name %in% input$selected_descs) %>%
        mutate(name = factor(name, levels = input$selected_descs),
               age_years = age_months / 12, 
               ci.low = binom.confint(x = answer * n, n = n, tol = .1,
                                      methods = "bayes")$lower,
               ci.high = binom.confint(x = answer * n, n = n, tol = .1,
                                       methods = "bayes")$upper) %>%
        filter(n > MIN_N)
      
      label_data <- plot_data %>% 
        group_by(specific_category, name) %>% 
        summarise(answer = median(answer), 
                  age_years = median(age_years))
      
      ggplot(plot_data,
             aes(x = age_years, y = answer, col = name)) + 
        geom_point(aes(size = n)) + 
        geom_linerange(aes(ymin = ci.low, ymax = ci.high)) + 
        # geom_smooth(se=FALSE, aes(weight = n), method = "loess") + 
        geom_line() + 
        scale_size_continuous(guide= FALSE) + 
        scale_colour_manual(guide=FALSE, 
                            values = stable_order_palete(length(unique(input$selected_descs)))) +
        scale_fill_manual(guide=FALSE, 
                          values = stable_order_palete(length(unique(input$selected_descs)))) + 
        ylim(c(0,1)) + 
        xlim(c(0,2)) +
        xlab("Age (years)") + 
        ylab("Proportion parents responding yes") + 
        geom_label_repel(data = label_data, aes(fill = name, 
                                                label = name), 
                         force = 8, size = 3,
                         color = 'white',
                         box.padding = unit(0.25, "lines"),
                         point.padding = unit(0.5, "lines"))
    })
  },
  
  options = list(height = 700)
)
```

# WHO & CDI Milestones

Comparison to World Health Organization standards. These are only for locomotor milestones. 

```{r}
who <- ms %>%
  ungroup %>%
  filter(!is.na(WHO_months_50th_pcntl)) %>%
  mutate(name = factor(name, levels = unique(name)))

who_ms <- who %>%
  group_by(name) %>%
  summarise(who = mean(WHO_months_50th_pcntl),
            who_lower = mean(WHO_CI_lower), 
            who_upper = mean(WHO_CI_upper), 
            milestone = WHO_milestone[1], 
            x = 0, y = 1)

ggplot(filter(who, n > MIN_N), 
       aes(x = age_months, y = answer)) + 
  geom_point(aes(size = n)) + 
  geom_smooth() + 
  facet_wrap(~name) + 
  geom_vline(data = who_ms, aes(xintercept = who), lty = 2) + 
  xlim(c(0,24)) + ylim(c(0,1)) + 
  geom_text(data = who_ms, aes(x = x, y = y, 
                               label = milestone), 
            hjust = 0)
```

Some sample CDC milestones. 

```{r}
cdc <- ms %>%
  ungroup() %>%
  filter(!is.na(cdc_months)) %>%
  mutate(name = factor(name, levels = unique(name)))

cdc_ms <- cdc %>%
  group_by(name, cdc_category) %>%
  summarise(cdc = mean(cdc_months),
            wording = cdc_wording[1], 
            x = 0, y = 1)

cat <- "lang/comm"

ggplot(filter(cdc, n > MIN_N, cdc_category == cat), 
       aes(x = age_months, y = answer)) + 
  geom_point(aes(size = n)) + 
  geom_smooth(span = 1) + 
  facet_wrap(~name) + 
  geom_vline(data = filter(cdc_ms, cdc_category == cat), 
             aes(xintercept = cdc), lty = 2) + 
  xlim(c(0,24)) + ylim(c(0,1)) + 
  geom_text(data = filter(cdc_ms, cdc_category == cat), 
            aes(x = x, y = y, label = wording), 
            hjust = 0, size = 2) + 
  theme(strip.text = element_text(size = 6, hjust = 0))
```

## Interactive Browser for CDC categories

```{r, cache=FALSE}
shinyApp(
  ui = fluidPage(
    sidebarLayout(
      sidebarPanel(
        selectInput("cat", "CDC Category", 
                    selected = "movement", 
                    choices = as.character(unique(cdc$cdc_category)))
      ),
      mainPanel(
        plotOutput("catplot"))
    )
  ),
  
  server = function(input, output) {
    output$catplot <- renderPlot({
      
      ggplot(filter(cdc, n > MIN_N, cdc_category == input$cat), 
             aes(x = age_months, y = answer)) + 
        geom_point(aes(size = n)) + 
        geom_line(col = "blue") + 
        facet_wrap(~name) + 
        geom_vline(data = filter(cdc_ms, cdc_category == input$cat), 
                   aes(xintercept = cdc), lty = 2) + 
        xlim(c(0,24)) + ylim(c(0,1)) + 
        geom_text(data = filter(cdc_ms, cdc_category == input$cat), 
                  aes(x = x, y = y, label = wording), 
                  hjust = 0, size = 2)
    })
  }
)
```

# Developmental categories

## Locomotion

Here are the average trajectories for locomotor development. 

```{r}
locomotor_milestones <- c("crawling","walking","reflexes",
                          "sitting","standing","rolling")

plot_data <- filter(ms, 
                    specific_category %in% locomotor_milestones) %>%
  filter(n > 20) %>%
  group_by(age_months, specific_category) %>%
  summarise(answer = mean(answer)) %>%
  mutate(age_years = age_months / 12)

label_data <- plot_data %>% 
  group_by(specific_category) %>% 
  summarise(answer = min(answer), 
            age_years = min(age_years))

ggplot(plot_data,
       aes(x = age_years, y = answer, col = specific_category)) + 
  # geom_point() +
  # geom_line() +
  geom_smooth(se=FALSE, method = "loess") + 
  ylim(c(0,1)) + 
  xlim(c(0,2)) +
  xlab("Age (years)") + 
  ylab("Proportion parents responding yes") + 
  scale_colour_solarized(guide=FALSE) + 
  scale_fill_solarized(guide=FALSE) + 
  geom_label_repel(data = label_data, aes(fill = specific_category, 
                                          label = specific_category), 
                   size = 3,
                   color = 'white')
```

## Language

Selected language milestones - CDC milestone timing is shown along the bottom row. 

```{r}
selected_descs <- c("Smiles and begins to babble", 
                    "Points at an object that he/she wants",
                    "Begins to talk a bit (mama, dada)", 
                    "Says one to four words with meaning",
                    "Speaks with gestures",
                    "Combines words to form short phrases")

plot_data <- filter(ms, 
                    name %in% selected_descs) %>%
  filter(n > MIN_N) %>%
  mutate(name = factor(name, levels = selected_descs), 
         age_years = age_months / 12, 
         ci.low = binom::binom.confint(x = answer * n, n = n, 
                                       methods = "bayes")$lower,
         ci.high = binom::binom.confint(x = answer * n, n = n, 
                                        methods = "bayes")$upper)

label_data <- plot_data %>% 
  group_by(name, cdc_wording) %>% 
  summarise(answer = median(answer), 
            age_years = median(age_years), 
            cdc_years = mean(cdc_months, na.rm=TRUE)/12 + 
              rnorm(n = 1, mean = 0, sd = .05)) # to jitter

ggplot(plot_data,
       aes(x = age_years, y = answer, col = name)) + 
  geom_point(aes(size = n)) + 
  geom_linerange(aes(ymin = ci.low, ymax = ci.high)) + 
  # geom_smooth(se=FALSE, aes(weight = n), method = "loess") + 
  geom_line() + 
  scale_colour_manual(guide=FALSE, 
                      values = stable_order_palete(length(unique(selected_descs)))) +
  scale_fill_manual(guide=FALSE, 
                    values = stable_order_palete(length(unique(selected_descs)))) + 
  scale_size_continuous(guide= FALSE) + 
  ylim(c(0,1)) + 
  xlim(c(0,2.2)) +
  xlab("Age (years)") + 
  ylab("Proportion parents responding yes") + 
  geom_linerange(data = label_data, 
             aes(x = cdc_years, ymin = 0, 
                 ymax = .1, col = name), size = 2) + 
  geom_label_repel(data = label_data, aes(fill = name, 
                                          label = name), 
                   force = 8, size = 3,
                   color = 'white',
                   box.padding = unit(0.25, "lines"),
                   point.padding = unit(0.5, "lines")) 

```

