---
title: "Developmental milestone curves from parent report data"
author: "Mike & Emily"
date: "`r Sys.Date()`"
output: 
html_document:
  toc: true
  number_sections: true
---

```{r, echo=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=5, 
                      echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE)
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(langcog))
library(stringr)
library(tidyr)
library(ggrepel)
library(binom)
library(shiny)
theme_set(theme_bw())
```

```{r, echo=FALSE}
solarized_colors <- c("#268bd2", "#cb4b16", "#859900", "#993399", "#d33682",
                      "#b58900", "#2aa198", "#6c71c4", "#dc322f")

stable_order_palete <- function(num_values) {
  c(rep(solarized_colors, num_values %/% length(solarized_colors)),
    solarized_colors[1:(num_values %% length(solarized_colors))])
}
```

# Introduction

We are analyzing data from Kinedu, an app for parents of young children to do activities with their children. 

Note: some indicators have been dropped during milestone revision.

```{r}
load("data/kinedu_all_processed_3-11-16.RData")

skills <- read.csv("skills.csv") 
info <- read.csv("info.csv") %>%
  filter(!duplicated(name)) %>%
  left_join(skills) 

ms <- right_join(ms, skills) %>% # remove duplicated indicators
  mutate(name = eng_desc) %>%
  select(indicator_id, age_months, answer, n, skill, 
         general_category, name) 
```

# Developmental categories

## Locomotion

Here are the average trajectories for locomotor development. 

```{r}
locomotor_milestones <- c("Crawling","Early movements and coordination","Finger 
                          Dexterity", "Hand coordination","Head Control","Newborn
                          Reflexes and Posture", "Physical coordination", "Standing up",
                          "Turning and Sitting", "Walking")

plot_data <- filter(ms, 
                    skill %in% locomotor_milestones) %>%
  filter(n > 20) %>%
  group_by(age_months, skill) %>%
  summarise(answer = mean(answer)) %>%
  mutate(age_years = age_months / 12)

label_data <- plot_data %>% 
  group_by(skill) %>% 
  summarise(answer = min(answer), 
            age_years = min(age_years))

ggplot(plot_data,
       aes(x = age_years, y = answer, col = skill)) + 
  # geom_point() +
  # geom_line() +
  geom_smooth(se=FALSE, method = "loess") + 
  ylim(c(0,1)) + 
  xlim(c(0,2)) +
  xlab("Age (years)") + 
  ylab("Proportion parents responding yes") + 
  scale_colour_solarized(guide=FALSE) + 
  scale_fill_solarized(guide=FALSE) + 
  geom_label_repel(data = label_data, aes(fill = skill, 
                                          label = skill), 
             size = 3,
             color = 'white')
```

## Language

Here are the average trajectories for language development. 

```{r}
language_milestones <- c("Babbling", "Communicating through gestures", "First words", "Language comprehension", "Language production")

plot_data <- filter(ms, 
                    skill %in% language_milestones) %>%
  filter(n > 20) %>%
  group_by(age_months, skill) %>%
  summarise(answer = mean(answer)) %>%
  mutate(age_years = age_months / 12)

label_data <- plot_data %>% 
  group_by(skill) %>% 
  summarise(answer = min(answer), 
            age_years = min(age_years))

ggplot(plot_data,
       aes(x = age_years, y = answer, col = skill)) + 
  # geom_point() +
  # geom_line() +
  geom_smooth(se=FALSE, method = "loess") + 
  ylim(c(0,1)) + 
  xlim(c(0,2)) +
  xlab("Age (years)") + 
  ylab("Proportion parents responding yes") + 
  scale_colour_solarized(guide=FALSE) + 
  scale_fill_solarized(guide=FALSE) + 
  geom_label_repel(data = label_data, aes(fill = skill, 
                                          label = skill), 
             size = 3,
             color = 'white')
```


# Interactive exploration

Select a general category, a specific category, and then milestones from that category. 

```{r, cache=FALSE}

shinyApp(
  ui = fluidPage(
    sidebarLayout(
      sidebarPanel(
        selectInput("gen_cat", "General Category", 
                    selected = "Language and communication", 
                    choices = as.character(unique(ms$general_category))),
        uiOutput("spec_selector"),
        uiOutput("desc_selector")
      ),
      mainPanel(
        plotOutput("catplot"))
    )
  ),
  
  server = function(input, output) {
    
    specs <- reactive({
      as.character(unique(ms$skill[ms$general_category == input$gen_cat]))
    })
    
    descs <- reactive({
      as.character(unique(ms$name[ms$general_category == input$gen_cat &
                                    ms$skill == input$skill]))
    })
    
    output$spec_selector <- renderUI({
      selectInput("skill", "skill", 
                     choices = specs(), 
                     selected = specs()[1])
    })
    
    output$desc_selector <- renderUI({
      selectizeInput("selected_descs", "Labels", 
                     choices = descs(), 
                     selected = descs()[1],
                     options = list(maxItems = 9), 
                     multiple = TRUE)
    })
    
    output$catplot <- renderPlot({
      
      plot_data <- filter(ms, 
                          skill == input$skill, 
                          general_category == input$gen_cat,
                          name %in% input$selected_descs) %>%
        mutate(name = factor(name, levels = input$selected_descs),
               age_years = age_months / 12, 
               ci.low = binom.confint(x = answer * n, n = n, tol = .1,
                                             methods = "bayes")$lower,
               ci.high = binom.confint(x = answer * n, n = n, tol = .1,
                                              methods = "bayes")$upper) %>%
        filter(n > 50)
      
      label_data <- plot_data %>% 
        group_by(skill, name) %>% 
        summarise(answer = median(answer), 
                  age_years = median(age_years))
      
      ggplot(plot_data,
             aes(x = age_years, y = answer, col = name)) + 
        geom_point(aes(size = n)) + 
        geom_linerange(aes(ymin = ci.low, ymax = ci.high)) + 
        # geom_smooth(se=FALSE, aes(weight = n), method = "loess") + 
        geom_line() + 
        scale_size_continuous(guide= FALSE) + 
        scale_colour_manual(guide=FALSE, 
                            values = stable_order_palete(length(unique(input$selected_descs)))) +
        scale_fill_manual(guide=FALSE, 
                          values = stable_order_palete(length(unique(input$selected_descs)))) + 
      ylim(c(0,1)) + 
      xlim(c(0,2)) +
      xlab("Age (years)") + 
      ylab("Proportion parents responding yes") + 
      geom_label_repel(data = label_data, aes(fill = name, 
                                              label = name), 
                       force = 8, size = 3,
                       color = 'white',
                       box.padding = unit(0.25, "lines"),
                       point.padding = unit(0.5, "lines"))
    })
  },
  
  options = list(height = 700)
  )
```

# Try to make growth curves

Explore just a single skill for now. 

```{r}
load("data/kinedu_all_3-11-16.RData")

locomotor_ids <- filter(skills, skill %in% locomotor_milestones)

locomotor <- d %>%
  filter(indicator_id %in% locomotor_ids$indicator_id) %>%
  left_join(select(locomotor_ids, indicator_id, skill, level, milestone))
```

Construct locomotor indices.

```{r}
n_locomotor <- locomotor %>%
  group_by(baby_id, skill) %>%
  summarise(n = n())

qplot(n, data = filter(n_locomotor, n > 1), binwidth = 1) + 
  xlim(c(0,20))
```

Take 2k kids with more than 5 milestones.

```{r}
library(quantregGrowth)

lms <- locomotor %>%
  group_by(baby_id, age_months, skill) %>%
  summarise(prop = mean(answer), 
            n = n()) %>%
  filter(n > 10)

mod <- gcrq(formula = prop ~ ps(Age, monotone = 1, lambda = 1000), 
     tau = taus, data = cdidata)

ages <- 8:18
newdata <- data.frame(Age = ages)

preds <- predictQR_fixed(mod, newdata = newdata) %>%
  data.frame %>%
  mutate(age = ages) %>%
  gather(percentile, pred, starts_with("X")) %>%
  mutate(percentile = as.character(as.numeric(str_replace(percentile, "X", "")) * 100))


ggplot(filter(lms, skill == "Crawling"), 
       aes(x = age_months, y = prop)) + 
  geom_jitter() 
```