---
title: "Parenting Project - Database exploration"
author: "Mike"
date: "`r Sys.Date()`"
output: 
  html_document:
  toc: true
number_sections: true
runtime: shiny
---
  
TODO:

* geoip?
* newer participants zero out indicators prior to initiation
* check distribution of answers vs. tries
* filter to first try or last try? 
  
#  Preliminary database loading

```{r echo=FALSE}
rm(list=ls())
knitr::opts_chunk$set(fig.width=8, fig.height=5, 
                      echo=TRUE, warning=FALSE, message=FALSE, cache=FALSE)
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(langcog))
library(stringr)
library(tidyr)
library(lubridate)
library(directlabels)
theme_set(theme_bw())
```

# Initial connection and baby info

Connect to database.

```{r echo=FALSE}
pwd <- read.table("pwd.txt", stringsAsFactors = FALSE)
kinedu <- src_mysql(host = "db.kinedu.com", 
                 dbname = "kinedu_app", 
                 user = "reader", password = pwd$V1[1])
```

Get baby info.

```{r}
baby_info <- tbl(kinedu, "babies") %>% 
  select(id, birthday, gender) %>%
  collect %>%
  mutate(baby_id = id, 
         dob = ymd(birthday)) %>%
  select(-birthday, -id)
```

# Indicators

## Extract indicator info

Get indicators data.

```{r}
indicator_cats <- tbl(kinedu, "indicator_categories") %>%
  mutate(indicator_category_id = id) %>%
  collect %>%
  data.frame

indicator_cats$name <- sapply(indicator_cats$name, 
                                function(x) {str_split(x, ",")[[1]][1]})
indicator_cats$cat_name <- str_sub(sapply(indicator_cats$name, 
                                function(x) {str_split(x, ":")[[1]][2]}), 
                              start = 2, end = -2)

indicator_cats <- indicator_cats %>% 
  select(indicator_category_id, cat_name)
  
indicator_info <- tbl(kinedu, "indicators") %>%
  mutate(indicator_id = id) %>%
  select(indicator_id, description, indicator_category_id) %>%
  collect %>% 
  data.frame %>%
  left_join(indicator_cats)

indicators <- tbl(kinedu, "indicator_answers") 
```

Get short descriptions.

```{r}
indicator_info$eng_desc <- sapply(indicator_info$description, 
                                function(x) {str_split(x, ",")[[1]][1]})
indicator_info$eng_desc <- str_sub(sapply(indicator_info$eng_desc, 
                                function(x) {str_split(x, ":")[[1]][2]}), 
                              start = 2, end = -2)
indicator_info$short_desc <- str_sub(indicator_info$eng_desc, start = 1, end = 20)
```

## Pull indicators

Explore indicator answers. Use raw SQL to get a limited number of these for exploration. 

```{r}
indicators_local <- tbl(kinedu, 
                        sql("SELECT indicator_id, tried_count, baby_id, answer, created_at FROM indicator_answers LIMIT 100000")) %>%
  collect
```

What's the tried count and answer mean? 

```{r}
indicators_local %>%
  group_by(tried_count) %>%
  summarise(answer = mean(answer, na.rm=TRUE),
            prop.na = mean(is.na(answer)),
            n = n())
```

Merge in baby info. 

```{r}
d <- left_join(indicators_local, baby_info) %>%
  mutate(indicator_date = ymd_hms(created_at)) %>%
  filter(!is.na(gender)) %>%
  mutate(age_months = floor(as.numeric(difftime(indicator_date, 
                                         dob, units = "days"))/30.3)) %>%
  filter(tried_count == 0, age_months > 0, age_months < 24) 
```

Now get the most frequent of these, and use these to filter out the curves we want to plot. 

```{r}
freq_indicators <- d %>%
  group_by(indicator_id) %>%
  summarise(n = n()) %>%
  filter(n > quantile(n, .70))
  
ms <- d %>%
  filter(indicator_id %in% freq_indicators$indicator_id) %>%
  group_by(indicator_id, age_months) %>%
  summarise(answer = mean(answer)) %>%
  left_join(indicator_info) 
```

Plot.

```{r}
library(shiny)
shinyApp(
  ui = fluidPage(
    selectInput("cat_name", "Category:", 
                choices = unique(ms$cat_name)),
    plotOutput("catplot")
  ),
  
  server = function(input, output) {
    output$catplot <- renderPlot({
      ggplot(filter(ms,cat_name == input$cat_name),
             aes(x = age_months, y = answer, col = eng_desc)) + 
        geom_point() + 
        geom_smooth(se=FALSE) + 
        scale_colour_solarized() + 
        ylim(c(0,1))
    })
  },
  
  options = list(height = 700)
)
```