---
title: "Parenting Project - Database exploration"
author: "Mike"
date: "`r Sys.Date()`"
output: 
  html_document:
  toc: true
number_sections: true
runtime: shiny
---
  
TODO:

* geoip?
* newer participants zero out indicators prior to initiation
* check distribution of answers vs. tries
* filter to first try or last try? 
  
#  Preliminary database loading

```{r echo=FALSE}
rm(list=ls())
knitr::opts_chunk$set(fig.width=8, fig.height=5, 
                      echo=TRUE, warning=FALSE, message=FALSE, cache=FALSE)
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(langcog))
library(stringr)
library(tidyr)
library(lubridate)
library(directlabels)
theme_set(theme_bw())
```

# Initial connection and baby info

Connect to database.

```{r echo=FALSE}
pwd <- read.table("pwd.txt", stringsAsFactors = FALSE)
kinedu <- src_mysql(host = "db-slave.kinedu.com", 
                 dbname = "kinedu_app", 
                 user = "reader", password = pwd$V1[1])
```

Get baby info.

```{r}
baby_info <- tbl(kinedu, "babies") %>% 
  select(id, birthday, gender) %>%
  collect %>%
  mutate(baby_id = id, 
         dob = ymd(birthday)) %>%
  select(-birthday, -id)
```

# Indicators

## Extract indicator info

Get indicators data.

```{r}
indicator_cats <- tbl(kinedu, "indicator_categories") %>%
  mutate(indicator_category_id = id) %>%
  collect %>%
  data.frame

indicator_cats$name <- sapply(indicator_cats$name, 
                                function(x) {str_split(x, ",")[[1]][1]})
indicator_cats$cat_name <- str_sub(sapply(indicator_cats$name, 
                                function(x) {str_split(x, ":")[[1]][2]}), 
                              start = 2, end = -2)

indicator_cats <- indicator_cats %>% 
  select(indicator_category_id, cat_name)
  
indicator_info <- tbl(kinedu, "indicators") %>%
  mutate(indicator_id = id) %>%
  select(indicator_id, description, indicator_category_id) %>%
  collect %>% 
  data.frame %>%
  left_join(indicator_cats)
```

Get short descriptions.

```{r}
indicator_info$eng_desc <- sapply(indicator_info$description, 
                                function(x) {str_split(x, ",")[[1]][1]})
indicator_info$eng_desc <- str_sub(sapply(indicator_info$eng_desc, 
                                function(x) {str_split(x, ":")[[1]][2]}), 
                              start = 2, end = -2)
indicator_info$short_desc <- str_sub(indicator_info$eng_desc, start = 1, end = 20)
```

## Pull indicators

Explore indicator answers. Use raw SQL to get a limited number of these for exploration. 

From Guillermo:

> `SELECT SUM(tried_count + 1), indicator_id, MAX(answer) as max_answer FROM indicator_answers GROUP BY indicator_id HAVING max_answer = 1;`

add `LIMIT 1000000` to get only the first million.

```{r}
indicators <- tbl(kinedu, "indicator_answers") 

indicators_local <- tbl(kinedu, 
                        sql("SELECT indicator_id, tried_count, baby_id, answer, created_at FROM indicator_answers")) %>%
  collect
```

Merge in baby info. 

Take only questions that have been tried once. 

```{r}
d <- left_join(indicators_local, baby_info) %>%
  mutate(indicator_date = ymd_hms(created_at)) %>%
  filter(!is.na(gender)) %>%
  mutate(age_months = floor(as.numeric(difftime(indicator_date, 
                                         dob, units = "days"))/30.3)) %>%
  filter(tried_count == 0, age_months > 0, age_months < 24) 
```

Save to an RData file for quick loading later. 

```{r}
save(file = "kinedu_all_3-11-16.RData", list = c("indicator_info","d"))
```

Now get the most frequent of these, and use these to filter out the curves we want to plot. 

```{r}
freq_indicators <- d %>%
  group_by(indicator_id) %>%
  summarise(n = n()) 

ms <- d %>%
  filter(indicator_id %in% freq_indicators$indicator_id) %>%
  group_by(indicator_id, age_months) %>%
  summarise(answer = mean(answer), 
            n = n()) %>%
  left_join(indicator_info)
```

Now resave. 

```{r}
save(file = "kinedu_all_processed_3-11-16.RData", list = c("ms"))
```